LEGWORK: A python package for computing the evolution and detectability \\ of stellar-origin gravitational-wave sources with space-based detectors

### 2022-01-25 src/ms.tex
it should be frequency space
average the final SNRs too

\newcommand{\todo}[1]{{\color{red}{[TODO: #1}]}}
\newcommand{\referee}[1]{{\color{WildStrawberry} #1}}

@@

\begin{equation}{}
\label{eq:signal}
    \begin{split}
        |\tilde{s}(f)|^2 = |&F_+(\theta, \phi, \psi)\tilde{h}_+(t, D_L, \iota, \beta) \\
        &+ F_{\times}(\theta, \phi, \psi)\tilde{h}_{\times}(t, D_L, \iota, \beta)|^2,
    \end{split}
\end{equation}
where $F_{+,\times}$ are the `plus' and `cross' antenna patterns of the LISA detector to the `plus' and `cross' strains, $h_{+,\times}$. Note throughout any parameters discussed with the subscript $x_{+,\times}$ refers to both $x_{+}$ and $x_{\times}$.

In LISA's case, when averaged over all angles and polarisations, the antenna patterns are orthogonal and thus $\langle F_+ F_{\times}\rangle = 0$. This means we can rewrite Eq.~\ref{eq:signal} as 
\begin{equation}{}
    \begin{split}
        |\tilde{s}(f)|^2 &= |F_+(\theta, \phi, \psi)\tilde{h}_+(t, D_L, \iota, \beta)|^2\\
        &+ |F_{\times}(\theta, \phi, \psi)\tilde{h}_{\times}(t, D_L, \iota, \beta)|^2,
    \end{split}
\end{equation} 
which can then be applied to Eq.~\ref{eq:snr_general_start} as
\begin{equation}{}
\label{eq:snr_general_simpler}
    \rho^2 = 4 \int_0^{\infty} \frac{|F_+\tilde{h}_+|^2 + |F_{\times}\tilde{h}_{\times}|^2}{P_{\rm n}(f)} df.
\end{equation}

@@

For a evolving and circular binary, the sum can be removed and so the SNR found as
\begin{equation}{\lwFuncURL{snr}{snr_circ_evolving}}
    \rho^2_{\rm c, e} = \frac{1}{4} \int_{f_{2, i}}^{f_{2, f}} \frac{h_{c, 2}^2}{f_2^2 S_{\rm n}(f_2)} \dd{f_2}. \label{eq:snr_chirp_circ}
\end{equation}

For a stationary and eccentric binary we can approximate the integral.
\begin{align}
    \rho^2_{\rm e, s} &= \frac{1}{4} \sum_{n=1}^{\infty}  \lim_{\Delta f \to 0} \int_{f_{n}}^{f_{n} + \Delta f_n} \frac{h_{c, n}^2}{f_n^2 S_{\rm n}(f_n)} \dd{f_n}, \\
    &= \frac{1}{4} \sum_{n=1}^{\infty} \frac{\Delta f_n \cdot h_{c, n}^2}{f_n^2 S_{\rm n}(f_n)}, \\
    &= \frac{1}{4} \sum_{n=1}^{\infty} \frac{\dot{f}_n \Delta T \cdot h_{c, n}^2}{f_n^2 S_{\rm n}(f_n)}, \\
    &= \frac{1}{4} \sum_{n=1}^{\infty} \qty(\frac{\dot{f}_n}{f_n^2} h_{c, n}^2) \frac{T_{\rm obs}}{S_{\rm n}(f_n)},
\end{align}
where we have applied \eqref{eq:strain-charstrain} to convert between strains. This gives the following expression
\begin{equation}{\lwFuncURL{snr}{snr_ecc_stationary}}
    \rho^2_{\rm e, s} = \frac{1}{4} \sum_{n=1}^{\infty} \frac{h_{n}^2 T_{\rm obs}}{S_{\rm n}(f_n)}. \label{eq:snr_stat_ecc}
\end{equation}
Finally, for a stationary and circular binary the signal-to-noise ratio is
\begin{equation}{\lwFuncURL{snr}{snr_circ_stationary}}
    \rho^2_{\rm c, s} = \frac{1}{4} \frac{h_2^2 T_{\rm obs}}{S_{\rm n}(f_2)}.
    \label{eq:snr_stat_circ}
\end{equation}

@@

The most fundamental use case of \lw{} is to compute the SNR of an individual binary system. This can be accomplished in \lw{} with only two lines of code - one line to set up the source and another to compute the SNR. As an example one could consider a binary with the parameters
\begin{align*}
    m_1 = m_2 &= 10 \unit{M_{\odot}},\ d = 8 \unit{kpc},\\
    f_{\rm orb} &= 10^{-4} \unit{Hz},\ e = 0.2,
\end{align*}
where $m_1, m_2$ are the primary and secondary masses, $d$ is the distance to the source, $f_{\rm orb}$ is the orbital frequency and $e$ is the eccentricity. As we do not specify a position, polarisation or inclination, \lw{} will calculate the SNR averaged over these quantities. One can now instantiate a source in \lw{} using these parameters (leaving the mission parameters as the default values, such that we compute the SNR for a 4-year LISA mission). As shown in the linked demo, \lw{} quickly computes that the SNR of this binary is $4.49$. In the background, \lw{} decides which SNR approximation is most applicable given the eccentricity of the binary and whether it is stationary in frequency space.

>>>>

\newcommand{\todo}[1]{{\color{red}{[TODO: #1}]}}

% referee command so we can highlight changes
\definecolor{WildStrawberry}{RGB}{239, 85, 159}
\newcommand{\referee}[1]{{\color{WildStrawberry} #1}}

@@

\begin{equation}{}
\label{eq:signal}
    \begin{split}
        |\tilde{s}(f)|^2 = |&F_+(\theta, \phi, \psi)\tilde{h}_+(\referee{f}, D_L, \iota, \beta) \\
        &+ F_{\times}(\theta, \phi, \psi)\tilde{h}_{\times}(\referee{f}, D_L, \iota, \beta)|^2,
    \end{split}
\end{equation}
where $F_{+,\times}$ are the `plus' and `cross' antenna patterns of the LISA detector to the `plus' and `cross' strains, $h_{+,\times}$. Note throughout any parameters discussed with the subscript $x_{+,\times}$ refers to both $x_{+}$ and $x_{\times}$.

In LISA's case, when averaged over all angles and polarisations, the antenna patterns are orthogonal and thus $\langle F_+ F_{\times}\rangle = 0$. This means we can rewrite Eq.~\ref{eq:signal} as 
\begin{equation}{}
    \begin{split}
        |\tilde{s}(f)|^2 &= |F_+(\theta, \phi, \psi)\tilde{h}_+(\referee{f}, D_L, \iota, \beta)|^2\\
        &+ |F_{\times}(\theta, \phi, \psi)\tilde{h}_{\times}(\referee{f}, D_L, \iota, \beta)|^2,
    \end{split}
\end{equation} 
which can then be applied to Eq.~\ref{eq:snr_general_start} as
\begin{equation}{}
\label{eq:snr_general_simpler}
    \rho^2 = 4 \int_0^{\infty} \frac{|F_+\tilde{h}_+|^2 + |F_{\times}\tilde{h}_{\times}|^2}{P_{\rm n}(f)} df.
\end{equation}

@@

For a evolving and circular binary, the sum can be removed and so the SNR found as
\begin{equation}{\lwFuncURL{snr}{snr_circ_evolving}}
    \referee{\langle \rho \rangle^2_{\rm c, e}} = \frac{1}{4} \int_{f_{2, i}}^{f_{2, f}} \frac{h_{c, 2}^2}{f_2^2 S_{\rm n}(f_2)} \dd{f_2}. \label{eq:snr_chirp_circ}
\end{equation}

For a stationary and eccentric binary we can approximate the integral.
\begin{align}
    \referee{\langle \rho \rangle^2_{\rm e, s}} &= \frac{1}{4} \sum_{n=1}^{\infty}  \lim_{\Delta f \to 0} \int_{f_{n}}^{f_{n} + \Delta f_n} \frac{h_{c, n}^2}{f_n^2 S_{\rm n}(f_n)} \dd{f_n}, \\
    &= \frac{1}{4} \sum_{n=1}^{\infty} \frac{\Delta f_n \cdot h_{c, n}^2}{f_n^2 S_{\rm n}(f_n)}, \\
    &= \frac{1}{4} \sum_{n=1}^{\infty} \frac{\dot{f}_n \Delta T \cdot h_{c, n}^2}{f_n^2 S_{\rm n}(f_n)}, \\
    &= \frac{1}{4} \sum_{n=1}^{\infty} \qty(\frac{\dot{f}_n}{f_n^2} h_{c, n}^2) \frac{T_{\rm obs}}{S_{\rm n}(f_n)},
\end{align}
where we have applied \eqref{eq:strain-charstrain} to convert between strains. This gives the following expression
\begin{equation}{\lwFuncURL{snr}{snr_ecc_stationary}}
    \referee{\langle \rho \rangle^2_{\rm e, s}} = \frac{1}{4} \sum_{n=1}^{\infty} \frac{h_{n}^2 T_{\rm obs}}{S_{\rm n}(f_n)}. \label{eq:snr_stat_ecc}
\end{equation}
Finally, for a stationary and circular binary the signal-to-noise ratio is
\begin{equation}{\lwFuncURL{snr}{snr_circ_stationary}}
    \referee{\langle \rho \rangle^2_{\rm c, s}} = \frac{1}{4} \frac{h_2^2 T_{\rm obs}}{S_{\rm n}(f_2)}.
    \label{eq:snr_stat_circ}
\end{equation}

@@

The most fundamental use case of \lw{} is to compute the SNR of an individual binary system. This can be accomplished in \lw{} with only two lines of code - one line to set up the source and another to compute the SNR. As an example one could consider a binary with the parameters
\begin{align*}
    m_1 = m_2 &= 10 \unit{M_{\odot}},\ d = 8 \unit{kpc},\\
    f_{\rm orb} &= 10^{-4} \unit{Hz},\ e = 0.2,
\end{align*}
where $m_1, m_2$ are the primary and secondary masses, $d$ is the distance to the source, $f_{\rm orb}$ is the orbital frequency and $e$ is the eccentricity. As we do not specify a position, polarisation or inclination, \lw{} will calculate the SNR averaged over these quantities. \referee{If a user specifies the position, then if either of the polarisation or inclination are not specified then they will be randomly generated. LEGWORK does not allow users to specify polarisations or inclinations without positions as presume that the user must know the position if they know the polarisation or inclination of the source.} One can now instantiate a source in \lw{} using these parameters (leaving the mission parameters as the default values, such that we compute the SNR for a 4-year LISA mission). As shown in the linked demo, \lw{} quickly computes that the SNR of this binary is $4.49$. In the background, \lw{} decides which SNR approximation is most applicable given the eccentricity of the binary and whether it is stationary in frequency space.
